### 2015-07-09 narrative/data-description.tex
spelling, etc.

\subsection{Data Description}
\label{sec:data-description}
In order for the middleware and storage layer to make decisions
that are appropriate given the user's intentions, data must be
described in a way that communicates these intentions to
the system. In this project, we label the smallest unit of data
saved in the storage system as a \textit{chunk}.  A chunk consists of not only
raw data, but also metadata that expresses additional
knowledge about that data.
A user space variable, such as \textit{temperature}, may
consist of a collection of chunks, possibly with different accuracy or
resolution, and each chunk representing
a different portion of the variable.
A key advantage of having attributes at this lowest level is that it allows data
semantics, user intentions, QoS requirements, and the relationships between user
data to be readily captured and embedded with the raw data so that these
smallest pieces can each be managed independently or collectively, depending on
the need. As an example, retrieving and visualizing 3D field data generated from a simulation
requires the mesh (consiting of coordinates and connectivity variables) be
accessed simultaneously and with low latency to achieve a good user experience.
Managing data at this low level
bridges the semantics gap between applications, middleware, and storage
systems and allows the system to understand user-level data, execute QoS
requirements and policies, and optimize application and system performance.

%The new APIs will enable the bridge by specifying selectable
%performance/quality/- cost tradeoffs from both the application and system
%perspectives based upon the user guided rules/policy and runtime system
%monitoring status. It allows the middleware to make best possible decisions
%from the feedback of storage system knowledge, such that it will embed user
%intentions and the available system storage.  We want the system to give the
%users a certain amount of currency in terms of bandwidth, storage space on
%each level, and latency expectations. These notions will be fuzzy but they
%will allow the user to make ad-hoc decisions to figure out what needs to be
%saved.

@@

An important research question which me must address is the semantics of
describing the utility function, and the possibility that the users will
accept this, and if the system can then use this information, and make 
decisions based on this utility function. We realize that the system will
try to optimize across all of the different user request, and the utility
function must have defaults that users can inquire into, and change these
when they see that these defaults can have a possibile negative impact on
their knowledge discovery.

The assumptions which we will make in choosing a utility function is that:
1) There will only be a small number of priority levels that users will use
to re-prioritze their data; 2) The number of tiers in the storage system will
also be small (e.g. NVRAM, PFS, Campaign Storage, HPSS), 3) The default
lifetime of objects in each tier will be specified by the resource by
a tuple (size, lifetime).  The user can then utilize this information to
place information to the system such as: store the highest priority data
as high up on the hiearchy for 6 months, and then migrate this data to tape.
The user can also define the utility for CR files as that they have high
priority but their lifetime is for a short period of time; i.e. when the
next CR file appears on the system or the next two CR files. Research in
describing this utility function will be in conjunction with our users
and we will evaluate how policies can be placed in the system and how well
this utility function can be observed by the user.

%
\begin{wrapfigure}{r}{0.6\textwidth}
        \vspace{-2ex}
        \begin{centering} 
	\includegraphics[scale=0.8]{graphics/SSIO-bucket.pdf}
        %\vspace{1ex}
        \caption{The system manages the data across the various stages of the data lifecycle}
        \label{fig:ssio-bucket}
        \end{centering}
       \vspace{-1ex}
\end{wrapfigure}

>>>>

\subsection{Data Description}
\label{sec:data-description}
In order for the middleware and storage layer to make decisions
that are appropriate given the user's intentions, data must be
described in a way that communicates these intentions to
the system. In this project, we label the smallest unit of data
saved in the storage system as a \textit{chunk}.  A chunk consists of not only
raw data, but also metadata that expresses additional
knowledge about that data.
A user space variable, such as \textit{temperature}, may
consist of a collection of chunks, possibly with different accuracy or
resolution, and each chunk representing
a different portion of the variable.
A key advantage of having attributes at this lowest level is that it allows data
semantics, user intentions, QoS requirements, and the relationships between user
data to be readily captured and embedded with the raw data so that these
smallest pieces can each be managed independently or collectively, depending on
the need. As an example, retrieving and visualizing 3D field data generated from a simulation
requires the mesh (consisting of coordinates and connectivity variables) be
accessed simultaneously and with low latency to achieve a good user experience.
Managing data at this low level
bridges the semantics gap between applications, middleware, and storage
systems and allows the system to understand user-level data, execute QoS
requirements and policies, and optimize application and system performance.

%The new APIs will enable the bridge by specifying selectable
%performance/quality/- cost trade-offs from both the application and system
%perspectives based upon the user guided rules/policy and runtime system
%monitoring status. It allows the middleware to make best possible decisions
%from the feedback of storage system knowledge, such that it will embed user
%intentions and the available system storage.  We want the system to give the
%users a certain amount of currency in terms of bandwidth, storage space on
%each level, and latency expectations. These notions will be fuzzy but they
%will allow the user to make ad-hoc decisions to figure out what needs to be
%saved.

@@

An important research question which me must address is the semantics of
describing the utility function, and the possibility that the users will
accept this, and if the system can then use this information, and make 
decisions based on this utility function. We realize that the system will
try to optimize across all of the different user request, and the utility
function must have defaults that users can inquire into, and change these
when they see that these defaults can have a possible negative impact on
their knowledge discovery.

\begin{wrapfigure}{r}{0.6\textwidth}
        \vspace{-6ex}
        \begin{centering} 
	\includegraphics[scale=0.8]{graphics/SSIO-bucket.pdf}
        %\vspace{1ex}
        \caption{The system manages the data across the various stages of the data lifecycle}
        \label{fig:ssio-bucket}
        \end{centering}
      % \vspace{-1ex}
\end{wrapfigure}

The assumptions which we will make in choosing a utility function is that:
1) There will only be a small number of priority levels that users will use
to re-prioritize their data; 2) The number of tiers in the storage system will
also be small (e.g. NVRAM, PFS, Campaign Storage, HPSS), 3) The default
lifetime of objects in each tier will be specified by the resource by
a tuple (size, lifetime).  The user can then utilize this information to
place information to the system such as: store the highest priority data
as high up on the hierarchy for 6 months, and then migrate this data to tape.
The user can also define the utility for CR files as that they have high
priority but their lifetime is for a short period of time; i.e. when the
next CR file appears on the system or the next two CR files. Research in
describing this utility function will be in conjunction with our users
and we will evaluate how policies can be placed in the system and how well
this utility function can be observed by the user.

%

### 2015-07-09 narrative/placement.tex
spelling, etc.

The application workflows targeted in this proposal generate very large
amounts of data, which needs to be processed and analyzed before
potential insights from the simulations can be realized. 
The overall data life cycle can be viewed as consisting of four key phases. In the first phase data 
is generated by a simulation or through a data acquisition system,
in the case of experimental and observational data (EOD). This data has 
to be appropriately initially placed (and possibly partitioned) within the heterogenous 
multi-tier storage to optimize application I/O overheads.
%In our proposed heterogenous multi-tier storage platform, the system must
%decide where to place the output data at this time, and how to partition and
%reorganize the data to optimize this initial placement. The key optimization
%metrics during this phase is the perceived I/O overhead for the application.

@@

\paragraph{Soliciting application {\em hints}:}
As described in \S\ref{sec:data-refactor} one of the main components of our
proposed storage system is the ability to refactor and reduce data as it is
generated, and to reorganize and regenerate the data as it is accessed. We
carry this principle into the placement and movement of data by allowing
applications to define hints and policies that guide \textbf{what} data is
placed \textbf{where}. We will explore the use of application hints in two
distinct areas. Firstly, we will study the challenges and tradeoffs of either
augmenting the I/O interface with hints, or allowing the addition of an
external specification that defines the use case. Our experience with
developing modern I/O interfaces has shown that both techniques have value,
and we will investigate the set of hints that are embedded in the
application code vs. those that are described within an non-compiled
specification. Secondly, we will study how hints can guide the initial data
placement as data is handed off from application to storage (during a
write) and from storage to application (during the read). In both cases we
will study what minimal set of annotations and hints can allow the storage
system to minimize data movement and optimize the resources consumed by I/O.

@@

\paragraph{Background:} 
Once the system has completed the initial data placement with minimal
overhead to the application, it is up to the storage system to manage the
migration of data to other storage servers and across the hierarchy. An
important reserach aspect here is how a liberated Hierarchical Storage
Manager (HSM) designed for online use, such as Sirocco, might be employed in
providing support for the migration of data. Sirocco autonomously groups the
distributed set of media and service nodes into like groups and manages the
movement of data between these groups. The grouping is managed using
attributes such as latency to first byte, bandwidth to and from the media, and
some function representing the resilience capabilities of the attached media.
In addition to these static attributes, when grouping servers into
pseudo-tiers, Sirocco also may employ more dynamic attributes such as CPU load
and the rate-of-change of the media use.

@@

\paragraph{Bounded estimates for lookup:}
One challenge for storage systems with continous decentralized migration is
the timeliness of lookup when applications issue read calls. This is due to
how no central directory exists for where data is at anytime. There are
solutions to this problem that utilize lookup tables which are updated
whenever a server initiates a migration process. However, these solutions
are not scalable. In order to have predictable performance the lookup
process must complete in a bounded time. Another solution here is to use a
distributed hashing function to determine exactly where data is placed,
reducing lookup to an $O(1)$ operation. We explore the discovery of data in
\S~\ref{sec:naming-discovery}. To support this discovery process, we will
investiagte how the migration of data can be performed in a way that enables
bounded estimates about time to first byte for a potential read operation.

@@

\subsubsection{Making data avilable on multiple tiers}
\label{sec:manag-data-mult}

\paragraph{Background:} Providing data to users for knowledge discovery is
one of the primary purposes of our storage system design. There is a
spectrum of approaches on how
this can be accomplished: On one end the storage system can aim to maintain
full resolution data in the highest tiers of the storage hierarchy. This
approach is, obviously, limited due to the disparity between size of data
and available space in the higher tiers. Another approach is to utilize the
techniques described in \ref{sec:data-refactor} to create auxiliary views of
the data that, with user defined error, represent the data, and place these
smaller data sets on the higher tiers for faster access.

There are a
number of drivers that push approaches to one or the other extreme:
(1) if a high tier has sufficient room for an application's working
set, approaches that move the working set into that tier will perform
well; Conversely, if a high tier very rarely captures working sets,
it is better to fill it with auxiliary data that speeds up access
to the working set captured by a lower tier; working sets only
reduce capacity misses but the concept can be generalized to the
likelihood of having the right data in place, i.e. proactively
moving data to high tiers accessed in predictable patterns reduces
compulsory misses as well; (2) if the application generates parallel
read/write requests and strong consistency semantics are important,
approaches that use auxiliary data to speed up access to persistent
shared data might yield better performance than approaches that
require coherence overhead to keep all copies on higher tiers
consistent; (3) if the application's access patterns are mismatched
with low tier access characteristics, a combination of approaches
that use auxiliary data and move data between tiers in order to
convert access patterns into a better match will perform well.

% Any other drivers?
\paragraph{Thinning of Data:} 
This spectrum enables interesting strategies
to manage space pressure on high tiers: instead of just evicting
pieces of data (blocks, pages, objects, files), one can free space
by ``thinning'' data so that accesses within more limited views or
lower resolutions can still occur without misses while requests
outside of views or higher resolutions can leverage the information
stored in auxiliary data. Conversely, data on a high tier can be
``enriched'' (e.g. turned into a higher resolution) if there is
value in doing so. We will explore how the storage system can perform these
operations to meet the service level objectives for each use case, while
optimizing the overall utility from the system perpective.

\paragraph{Locality Properties:} The decision on when to thin and when to
enrich data on a particular level depends on \emph{access patterns}.
Approaches that leverage access patterns in applications at leadership facilities
 have shown to greatly improve performance and reduce the amount
of overhead required for data management~\cite{he:hpdc13}. The
challenge is for applications to communicate these access patterns
to the I/O stack layers that manages tiers outside of the address
space. In the proposed work we will develop an abstraction for
\emph{locality} that allow applications or runtime profilers to
describe locality properties of access patterns, and implement
services that leverage locality properties to dynamically determine
the value of thinning or enriching data at a particular tier.
Consider the example of a road navigation system: the locality
properties of each navigation client are based on physical constraints
of speed, direction, likely resolution of map, and the fact that
clients almost never go offroad.

>>>>

The application workflows targeted in this proposal generate very large
amounts of data, which needs to be processed and analyzed before
potential insights from the simulations can be realized. 
The overall data life cycle can be viewed as consisting of four key phases. In the first phase data 
is generated by a simulation or through a data acquisition system,
in the case of experimental and observational data (EOD). This data has 
to be appropriately initially placed (and possibly partitioned) within the heterogeneous 
multi-tier storage to optimize application I/O overheads.
%In our proposed heterogenous multi-tier storage platform, the system must
%decide where to place the output data at this time, and how to partition and
%reorganize the data to optimize this initial placement. The key optimization
%metrics during this phase is the perceived I/O overhead for the application.

@@

\paragraph{Soliciting application {\em hints}:}
As described in \S\ref{sec:data-refactor} one of the main components of our
proposed storage system is the ability to refactor and reduce data as it is
generated, and to reorganize and regenerate the data as it is accessed. We
carry this principle into the placement and movement of data by allowing
applications to define hints and policies that guide \textbf{what} data is
placed \textbf{where}. We will explore the use of application hints in two
distinct areas. Firstly, we will study the challenges and trade-offs of either
augmenting the I/O interface with hints, or allowing the addition of an
external specification that defines the use case. Our experience with
developing modern I/O interfaces has shown that both techniques have value,
and we will investigate the set of hints that are embedded in the
application code vs. those that are described within an non-compiled
specification. Secondly, we will study how hints can guide the initial data
placement as data is handed off from application to storage (during a
write) and from storage to application (during the read). In both cases we
will study what minimal set of annotations and hints can allow the storage
system to minimize data movement and optimize the resources consumed by I/O.

@@

\paragraph{Background:} 
Once the system has completed the initial data placement with minimal
overhead to the application, it is up to the storage system to manage the
migration of data to other storage servers and across the hierarchy. An
important research aspect here is how a liberated Hierarchical Storage
Manager (HSM) designed for on-line use, such as Sirocco, might be employed in
providing support for the migration of data. Sirocco autonomously groups the
distributed set of media and service nodes into like groups and manages the
movement of data between these groups. The grouping is managed using
attributes such as latency to first byte, bandwidth to and from the media, and
some function representing the resilience capabilities of the attached media.
In addition to these static attributes, when grouping servers into
pseudo-tiers, Sirocco also may employ more dynamic attributes such as CPU load
and the rate-of-change of the media use.

@@

\paragraph{Bounded estimates for lookup:}
One challenge for storage systems with continuous decentralized migration is
the timeliness of lookup when applications issue read calls. This is due to
how no central directory exists for where data is at anytime. There are
solutions to this problem that utilize lookup tables which are updated
whenever a server initiates a migration process. However, these solutions
are not scalable. In order to have predictable performance the lookup
process must complete in a bounded time. Another solution here is to use a
distributed hashing function to determine exactly where data is placed,
reducing lookup to an $O(1)$ operation. We explore the discovery of data in
\S~\ref{sec:naming-discovery}. To support this discovery process, we will
investigate how the migration of data can be performed in a way that enables
bounded estimates about time to first byte for a potential read operation.

@@

\subsubsection*{Making data available on multiple tiers}
\label{sec:manag-data-mult}

\paragraph{Background:} Providing data to users for knowledge discovery is
one of the primary purposes of our storage system design. There is a
spectrum of approaches on how
this can be accomplished: On one end the storage system can aim to maintain
full resolution data in the highest tiers of the storage hierarchy. This
approach is, obviously, limited due to the disparity between size of data
and available space in the higher tiers. Another approach is to utilize the
techniques described in \S\ref{sec:data-refactor} to create auxiliary views of
the data that, with user defined error, represent the data, and place these
smaller data sets on the higher tiers for faster access.

There are a
number of drivers that push approaches to one or the other extreme:
(1) if a high tier has sufficient room for an application's working
set, approaches that move the working set into that tier will perform
well; Conversely, if a high tier very rarely captures working sets,
it is better to fill it with auxiliary data that speeds up access
to the working set captured by a lower tier; working sets only
reduce capacity misses but the concept can be generalized to the
likelihood of having the right data in place, i.e. pro-actively
moving data to high tiers accessed in predictable patterns reduces
compulsory misses as well; (2) if the application generates parallel
read/write requests and strong consistency semantics are important,
approaches that use auxiliary data to speed up access to persistent
shared data might yield better performance than approaches that
require coherence overhead to keep all copies on higher tiers
consistent; (3) if the application's access patterns are mismatched
with low tier access characteristics, a combination of approaches
that use auxiliary data and move data between tiers in order to
convert access patterns into a better match will perform well.

% Any other drivers?
\paragraph{Thinning of Data:} 
This spectrum enables interesting strategies
to manage space pressure on high tiers: instead of just evicting
pieces of data (blocks, pages, objects, files), one can free space
by ``thinning'' data so that accesses within more limited views or
lower resolutions can still occur without misses while requests
outside of views or higher resolutions can leverage the information
stored in auxiliary data. Conversely, data on a high tier can be
``enriched'' (e.g. turned into a higher resolution) if there is
value in doing so. We will explore how the storage system can perform these
operations to meet the service level objectives for each use case, while
optimizing the overall utility from the system perspective.

\paragraph{Locality Properties:} The decision on when to thin and when to
enrich data on a particular level depends on \emph{access patterns}.
Approaches that leverage access patterns in applications at leadership facilities
 have shown to greatly improve performance and reduce the amount
of overhead required for data management~\cite{he:hpdc13}. The
challenge is for applications to communicate these access patterns
to the I/O stack layers that manages tiers outside of the address
space. In the proposed work we will develop an abstraction for
\emph{locality} that allow applications or runtime profilers to
describe locality properties of access patterns, and implement
services that leverage locality properties to dynamically determine
the value of thinning or enriching data at a particular tier.
Consider the example of a road navigation system: the locality
properties of each navigation client are based on physical constraints
of speed, direction, likely resolution of map, and the fact that
clients almost never go off-road.
